% !TeX program = pdflatex
\documentclass[12pt,a4paper]{article}

% ==== Base packages ====
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[top=2.5cm,bottom=2.5cm,left=2.5cm,right=2.5cm]{geometry}
\usepackage{microtype}
\usepackage{csquotes}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{float}
\usepackage{booktabs}
\usepackage{array}
\usepackage{longtable}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{listings}

% ==== Styles ====
\hypersetup{
  colorlinks=true,
  linkcolor=blue!50!black,
  urlcolor=blue!50!black,
  citecolor=blue!50!black,
  pdfauthor={Jackson J. Alcázar},
  pdftitle={User Guide and Technical Reference -- MLR-X 1.0},
  pdfcreator={MLR-X}
}

\lstdefinestyle{mlrx}{
  basicstyle=\ttfamily\small,
  frame=single,
  breaklines=true,
  showstringspaces=false,
  columns=fullflexible
}

\setlist{nosep,leftmargin=2.1em}
\onehalfspacing

\title{\vspace{-1.2cm}\textbf{MLR-X 1.0}\\User Guide and Technical Reference}
\author{Jackson J. Alcázar\\MLR-X Project}
\date{\today}

\begin{document}
\maketitle

\begin{center}
\textit{This document describes the operational use of the MLR-X desktop application for data preparation, multiple linear regression (MLR) model search, validation, diagnostics, visualization, reporting, and command-line (CLI) automation.}
\end{center}

\tableofcontents
\newpage

% =========================================================
\section{Identity and scope}
\begin{itemize}
  \item \textbf{Name and version}: MLR-X (version 1.0).
  \item \textbf{License}: GNU Affero General Public License v3.0 (AGPL-3.0) with 2025 rights held by Jackson J. Alcázar. Windows build metadata mirrors this information.
  \item \textbf{Audience}: users who need to fit, validate, and diagnose multiple linear regression models on small or large datasets.
  \item \textbf{Official channels}:\\
  \textbf{Repository}: \url{https://github.com/Jacksonalcazar/MLR-X}\\
  \textbf{Homepage}: \url{jacksonalcazar.github.io/MLR-X}
\end{itemize}

\subsection{Distributions and compatibility}
\begin{itemize}
  \item \textbf{Available builds}: self-contained binaries for Linux, macOS (GUI), Windows GUI, and Windows GUI+CLI.
  \item \textbf{System requirements}: 64-bit Windows, Linux, or macOS. No Python installation or scientific libraries are required; everything ships within the build.
  \item \textbf{Console version}: available in the CLI-enabled builds (Linux and Windows GUI+CLI) to run the full pipeline without the GUI by supplying a \texttt{config.conf} file.
\end{itemize}

% =========================================================
\section{Requirements and installation}
The distributed builds are self-contained downloads and do not require installers or shortcut generators. Place each executable in your preferred folder. Optionally, add that folder to your system \texttt{PATH} so you can launch \verb|MLR-X| from any directory (e.g., \verb|export PATH=/opt/MLRX:$PATH| on Linux or configuring environment variables on Windows). Choose the appropriate build and follow these steps:

\subsection{Setting up the portable builds}
\begin{itemize}
  \item \textbf{Linux}: download the binary, move it to a stable folder, and mark it as executable if needed (\texttt{chmod +x MLR-X}). To add it to your \texttt{PATH}, prepend the folder with \verb|export PATH=/path/to/folder:$PATH| or the equivalent shell profile setting. To register a desktop launcher:
        \begin{enumerate}[label=\arabic*.]
          \item Build the Linux executable (see the steps above), then copy the binary and the icon to a stable location, for example:
          \item \texttt{sudo mkdir -p /opt/MLRX}
          \item \texttt{sudo cp MLRX /opt/MLRX/}\\
                \texttt{sudo cp icon/icon.ico /opt/MLRX/}
          \item Update the \texttt{Exec} and \texttt{Icon} paths in \texttt{MLRX.desktop} if you choose a different installation directory.
          \item Install the launcher by copying it to your applications directory and mark the file as executable:
          \item \texttt{chmod +x MLRX.desktop}
          \item \texttt{cp MLRX.desktop ~/.local/share/applications/}
          \item \# or for all users:\\
                \texttt{\# sudo cp MLRX.desktop /usr/share/applications/}
          \item Refresh the desktop launcher database (optional but recommended):
          \item \texttt{update-desktop-database ~/.local/share/applications}\\
                \texttt{\# or sudo update-desktop-database /usr/share/applications}
        \end{enumerate}
  \item \textbf{Windows GUI}: download the GUI build, place \texttt{MLR-X.exe} in a convenient folder, and run it directly.
  \item \textbf{Windows GUI+CLI}: same as Windows GUI, with the Run CLI option enabled.
  \item \textbf{macOS GUI}: download the app bundle and run it from your preferred location (e.g., move it to \texttt{/Applications} if desired).
\end{itemize}

% =========================================================
\section{Command-line (CLI) execution}
Available only in the builds that include the CLI (Linux and Windows GUI+CLI). Run the pipeline headlessly with a configuration file:
\begin{lstlisting}[style=mlrx]
# Linux
./MLRX config.conf or ./MLRX path/to/config.conf
# Windows GUI+CLI
MLRX.exe config.conf or MLRX.exe path/to/config.conf
\end{lstlisting}
Executing with a configuration file loads the defined parameters and runs the full pipeline in the console.

% =========================================================
\section{Interface tour (GUI)}
MLR-X organizes its features into tabs within a main window.

\subsection{\textbf{Setup \& Run} tab}
\paragraph{Dataset}
\begin{enumerate}[label=\arabic*.]
  \item Select the CSV (or use \emph{Browse}).
  \item Define the delimiter (\verb|;| by default), the dependent variable, and non-predictor columns.
  \item Exclude observations by ID or range (e.g., \texttt{3,9,4-8}).
  \item Enable \emph{Exclude nearly constant predictors} and set the percentage threshold.
  \item Actions: \emph{Preview data}, \emph{Run analysis}, \emph{Cancel}. Status appears in the top-right corner of the frame; the task name appears in the bottom-right corner.
\end{enumerate}

\paragraph{Data splitting}
Choose how to partition the dataset for training and testing.
\begin{itemize}
  \item \textbf{None}: use the full dataset.
  \item \textbf{Random}: specify \emph{Test size (\%)}.
  \item \textbf{Manual}: lists or ranges for train/test without overlap.
\end{itemize}
If you select \textbf{None} (full dataset), you can still perform external validation later in the \textbf{Validation} tab by loading an external file.

\paragraph{Settings}
  \begin{enumerate}[label=\arabic*.]
   \item Search method: \textbf{EPR-S} (Expand--Perturb--Reduce--Swap) or \textbf{All Subsets}. EPR-S is a new, original method implemented in this software (doi.org/xxxxx).
   \item EPR-S parameters: \emph{Number of seeds}, \emph{Seed size}, and \emph{Set iterations/seed}. Seeds are the independent starting points used to explore model combinations. Recommendation buttons help populate these values. Set iterations/seed can run in \textit{auto} (estimate and apply the maximum per seed automatically), \textit{manual} (requires max iterations per seed), or \textit{converge (run until convergence)}, and this last option may take a long time.
  \item Key parameters: \emph{Max predictors per model}, \emph{Significance level}, \emph{Correlation threshold}, \emph{VIF threshold}, \emph{Minimum report R$^2$}, \emph{Top models to report}, and \emph{Parallel jobs}. \emph{Top models to report} caps the number of best-scoring models exported to CSV, diagnostics tables, visualizations, and summaries (default 5000). Correlation threshold and Variance Inflation Factor (VIF) threshold control multicollinearity pruning. Minimum report R$^2$ uses the training-set R$^2$ as a gatekeeper even when optimizing alternative metrics.
  \item The \emph{target metric} (R$^2$, adj-R$^2$, or Q$^2$) sets the scoring objective for model evaluation, while the \emph{covariance type} (Non-robust, HC0--HC3) selects the heteroskedasticity-consistent variance estimator for reliable inference. Non-robust maintains the standard OLS covariance matrix, valid only under homoskedasticity. HC0 implements the classic White (1980) estimator, replacing the constant error variance with squared residuals. HC1 applies a small-sample degrees-of-freedom adjustment to HC0. HC2 down-weights high-leverage observations through the diagonal of the hat matrix ($h_i$), using the weighting factor $1/(1-h_i)$. HC3 applies the most conservative leverage adjustment with weighting factor $1/(1-h_i)^2$, approximating the jackknife estimator and providing maximum protection against influential observations. The progression from HC0 to HC3 represents increasing robustness to both heteroskedasticity and leverage points in finite samples.
   \item \emph{Clip predictions to}: define lower and/or upper bounds to constrain predicted values for scoring and exports.
  \end{enumerate}

\subsection{\textbf{Models} tab}
Displays three synchronized tables summarizing training, internal validation, and external validation results. The upper panel allows loading a results file (\emph{Browse}, \emph{Load}, \emph{Clear}, \emph{Filter}), selecting the sorting criterion, and filtering models by predictors.
\begin{itemize}
  \item \textbf{Top models training}: shows each model together with its predictor set, number of predictors (N pred), goodness-of-fit statistics (R$^2$, adjusted R$^2$, RMSE, MAE, and $s$), and multicollinearity indicators (maximum and average VIF).
  \item \textbf{Top models internal validation (CV)}: reports the corresponding LOO and k-fold performance metrics, including Q$^2$, RMSE, MAE, and $s$ for each procedure.
  \item \textbf{Top models external validation}: presents the external prediction results using Q$^2_{\mathrm{F1}}$, Q$^2_{\mathrm{F2}}$ (R$^2_{\mathrm{ext}}$), Q$^2_{\mathrm{F3}}$, and the external RMSE and MAE values.
\end{itemize}
All three tables refresh automatically and remain synchronized with the Validation, Diagnostics, Visualization, and Summary tabs, which become available as soon as a model set is generated when the run analysis is completed or when model results are loaded.

\subsection{\textbf{Validation} tab}
Use this tab to configure cross-validation and external validation for the selected models.
\begin{itemize}
  \item Model selection: \emph{Top}, \emph{Multiple} (range like \texttt{1-5,7,9}), or \emph{All}.
  \item \textbf{LOO and k-fold}: internal validation with \emph{Leave One Out (LOO)} and/or \emph{Folds} and \emph{Repeats}.
  \item \textbf{External validation}: select file and delimiter; requires the same target column as the training set.
\end{itemize}
Validation results are automatically appended to the Models tab tables when each process completes.

\subsection{\textbf{Diagnostics} tab}
Provides per-observation diagnostics for the selected model and the corresponding correlation matrix.
\begin{itemize}
  \item Model selector with a status message for availability, plus a dataset filter to display \emph{Training only}, \emph{Training and testing}, or \emph{Testing only}; the holdout status propagates automatically.
  \item Observation diagnostics table showing \emph{Observation}, \emph{Set}, \emph{Actual}, \emph{Predicted}, \emph{Predicted (LOO)}, \emph{Residual}, \emph{Residual (LOO)}, \emph{Z value}, \emph{Leverage}, \emph{Std. residual}, and \emph{Std. residual by LOO}.
  \item \emph{Export CSV} exports the currently visible observation diagnostics using the active dataset filter.
  \item Variable correlation matrix (lower triangle) with a toggle to include the dependent variable when available, plus CSV export for the displayed matrix.
\end{itemize}

\subsection{\textbf{Visualization} tab}
\begin{itemize}
  \item Model selector with availability status and dataset filter (\emph{Training only}, \emph{Training and testing}, or \emph{Testing only}); Y-randomization analysis runs per model and can insert its charts into the plot list once computed.
  \item Chart list: correlation heatmaps (with or without the dependent variable), Predicted vs Actual (standard and by LOO), residuals vs actual/predictions (standard and by LOO), Scale--Location (standard and by LOO), QQ residuals (standard and by LOO), residual distribution (standard and by LOO), Williams plot (standard and by LOO), Cook's distance vs leverage (standard and by LOO), leverage plot, and Predicted vs Leverage.
  \item Styling and overlays: legend toggle with four preset positions, labeling mode (observations or an auto-detected extra column), identity line (for predicted vs actual and QQ residual plots), optional linear fit (Predicted vs Actual), $h^*$ threshold color for influence plots, grid toggle, and synchronized axis ranges.
  \item Marker customization: separate color and marker-style pickers for training and testing points, plus marker size control; heatmaps disable incompatible controls automatically.
  \item Axes and export: manual limits for X/Y, editable axis labels with font size, and high-resolution export of the current plot to PNG, TIFF, PDF, or SVG.
\end{itemize}

\subsection{\textbf{Summary} tab}
\begin{itemize}
  \item Model selector with availability status plus a \emph{Use recommended covariance} toggle (applies BP/White and leverage diagnostics before switching to the suggested estimator) and \emph{Export Summary...} to save the full text dossier. A scrollable text pane renders the summary for the current model.
  \item Overview section: regression type, chosen covariance, search method, iterations per seed, target metric, models explored/found/reported (with threshold), significance level, observation counts, dependent variable, predictor list, k-fold plan, CPU times, clip-prediction bounds, and Y-randomization status.
  \item Covariance diagnostics (when the recommendation is enabled): recommended/applied type, Breusch--Pagan and White $p$-values, leverage threshold, maximum leverage, high-leverage count, and reason/notes that triggered the robust choice.
  \item Coefficients and equation: coefficient table with standard errors, $t$, $p$, and confidence intervals; equations rendered with and without errors.
  \item Metrics tables: grouped training metrics (R$^2$, adj-R$^2$, RMSE, MAE, $s$, VIF max/avg, $|r|_{\max}$), internal validation (LOO and k-fold blocks with repeat note), and external validation (Q$^2_{\mathrm{F1}}$, Q$^2_{\mathrm{F2}}$/R$^2_{\mathrm{ext}}$, Q$^2_{\mathrm{F3}}$, RMSE, MAE).
\end{itemize}

\subsection{\textbf{Desktop} tab (Variable Explorer)}
\begin{itemize}
  \item Dataset and results inputs: browse/load the CSV and an optional results file (unlocks model-based options) with a status line indicating readiness.
  \item Variable scope and axes: pick \emph{All variables} or \emph{Only variables of model} (enter an ID), then select X/Y axes. A \emph{Histogram} toggle switches the plot mode, with bin controls for a fixed number of bins or custom breakpoints.
  \item Plot controls: marker size/color/style, legend with selectable positions, gridlines, linear fit with optional R$^2$ display, partial residuals (when a model context is active), aspect ratio presets (1, 4:3, 16:9, 9:16), manual axis limits and tick steps, and a dialog to set axis labels and font size. Interactive labels can be added/dragged with left-click and cleared with right-click.
  \item Exports and filtered datasets: \emph{Save current plot} to PNG, TIFF, PDF, or SVG (high resolution by default). \emph{Create filtered dataset} builds a CSV containing observation/ID/non-variable columns plus predictors from the current model or a specified model range, warning about any missing models in the requested range.
\end{itemize}

\subsection{\textbf{Help} tab}
Summarizes product information and support links for the application. The About panel shows the MLR-X name, version, license notice, and tagline. Quick links open the homepage and contact email, and buttons provide access to the bundled manual, citation dialog, bug-report form, and donation page.

% =========================================================
\section{Analysis flow and process handling}
\subsection{Before starting}
\begin{enumerate}
  \item Load and preview the dataset, confirming delimiter, target, exclusions, and split mode.
  \item Check that no other run is in progress, then choose where to save results when prompted (defaults to \texttt{models.csv}; a custom path is allowed).
\end{enumerate}

\subsection{Starting a run}
\begin{enumerate}
  \item Click \emph{Run analysis}. The app validates the configuration, reloads the dataset (including the split), and prepares the selected output file.
  \item The status switches to \emph{Running}, the \emph{Run analysis} button is disabled, \emph{Cancel} is enabled, and previous results, caches, and downstream tabs (Validation, Summary, Diagnostics, Visualization, Desktop) are reset.
  \item A run header plus ``Starting analysis...'' is appended to the log; the progress bar and step counter are reset for the search and post-processing stages.
\end{enumerate}

\subsection{Progress and cancellation}
\begin{itemize}
  \item The progress bar tracks completion across stages, and the \emph{Log} shows model counts, filters, CPU times, and other events in real time.
  \item \emph{Cancel} changes the status to \emph{Cancelling...}, disables the \emph{Cancel} button, signals all workers to stop, and records the request in the log. Errors raise a dialog, set status to \emph{Error}, and keep the run controls enabled for a new attempt.
\end{itemize}

\subsection{Completion and outputs}
\begin{enumerate}
  \item When finished, the status becomes \emph{Completed}; \emph{Run analysis} re-enables and \emph{Cancel} disables. Validation, Diagnostics, Visualization, and Summary tabs are unlocked with the new results, and any external holdout is refreshed if available.
  \item Results load automatically into the tables, summary pane, and downstream tabs, and the CSV is written to the confirmed path (\texttt{models.csv} by default). The header line embeds run metadata as JSON between \verb|#EPRS-S ... ;#|, and the most recent export path is remembered for the next session.
  \item If no models meet the reporting threshold, the app clears tables, keeps analysis tabs disabled, and notifies you with an information dialog.
\end{enumerate}

% =========================================================
\section{Search algorithms}
\subsection{EPR-S (Expand--Perturb--Reduce--Swap)}
Heuristic search that runs per seed using the target metric, predictor cap, correlation threshold, VIF veto, and reporting thresholds defined in Setup \& Run. Each seed follows a deterministic loop:
\begin{enumerate}
  \item Initialize with the seed's starting predictors and compute the baseline metric.
  \item \textbf{Expand}: add the single predictor that most improves the metric without violating the maximum predictors.
  \item \textbf{Perturb}: perform one-for-one swaps (drop one, add one) to escape local optima.
  \item \textbf{Reduce}: prune one predictor at a time when the metric stays the same or improves.
  \item \textbf{Correlation clean-up}: if two active predictors exceed the correlation threshold, drop the one with weaker contribution (or lower target correlation when needed) and refresh the remaining set.
  \item \textbf{Swap recovery}: revisit previously dropped variables when swap logs suggest a possible gain.
  \item Stop when no further gains appear after repeated passes or when the iteration budget per seed is reached. Quality filters (VIF and minimum report R$^2$) keep only unique, eligible models sorted by the target metric.
\end{enumerate}

\subsection{All subsets (exhaustive)}
Enumerates every predictor combination up to \emph{Max predictors}, pre-filtering by the correlation threshold and applying VIF checks during evaluation. Progress tracking, cancellation, and the same export limits and reporting thresholds as EPR-S are enforced. Use this engine when the predictor space is small (practical below roughly 25 variables) because the number of combinations grows quickly.

\subsection{Engine good practices}
\begin{itemize}
  \item Seeds: at least as many as predictors; use \emph{Recommended} to auto-populate sensible values.
  \item Seed size: about half of \emph{Max predictors} balances exploration and feasibility.
  \item VIF threshold: keep it $<5$ to curb multicollinearity; tune correlation threshold between 0 and 1 alongside the chosen significance level.
  \item Prefer EPR-S for large predictor spaces; reserve all subsets for small designs or confirmatory sweeps.
\end{itemize}

% =========================================================
\section{Result format and exports}
\subsection{Model CSV}
\begin{itemize}
  \item Semicolon-separated file with columns for model ID, predictors, predictor count, training metrics (R$^2$, adjusted R$^2$, RMSE, MAE, $s$), and multicollinearity indicators (maximum and average VIF).
  \item Internal validation columns append LOO and k-fold metrics (Q$^2$, RMSE, MAE, $s$); external validation columns append Q$^2$F1, Q$^2$F2 (external R$^2$), Q$^2$F3, RMSE, $s$, and MAE.
  \item The first row stores a serialized metadata block (dataset path, delimiter, dependent/ignored columns, constant-filter threshold, exclusions, target metric, search method, covariance type, iteration mode, split configuration, clipping limits, k-fold settings, validation file/delimiter, export limit, models found/explored/reported, and CPU times) so the UI can reload the configuration accurately.
  \item \emph{Export limit} caps the number of rows written; the exported table follows the training-order sort and mirrors the Models tab, including validation tables when available.
\end{itemize}

\subsection{Summary and reports}
\begin{itemize}
  \item Summary tab: \emph{Export Summary...} writes the currently selected model dossier (including the chosen covariance setting) to a text file.
  \item Diagnostics tab: \emph{Export CSV} saves the filtered observation diagnostics and the active correlation matrix (with or without the target) using semicolon-separated CSV.
  \item Visualization/plots: \emph{Save current plot} exports the visible chart to PNG, TIFF, PDF, or SVG with high-DPI output when needed.
\end{itemize}

% =========================================================
\section{Command-line usage (CLI)}
Available in the Linux and Windows GUI+CLI builds to run the full pipeline without launching the GUI. Provide exactly one argument: the path to a configuration file.

\subsection{Invocation}
\begin{lstlisting}[style=mlrx]
# Linux
./MLRX path/to/config.conf
# Windows GUI+CLI
MLRX.exe path/to/config.conf
\end{lstlisting}
If additional arguments are supplied, the program shows the expected syntax and exits.

\subsection{Configuration file structure}
Plain-text file with headed blocks (\# Dataset, \# Method, \# Validation, \# Output, etc.) and key--value pairs. List-type options let you select by index using \texttt{selected option: <n>} inside the block (for example, to pick a delimiter or target metric).

\paragraph{Main fields}
\begin{longtable}{>{\bfseries}p{4.2cm}p{10.2cm}}
data\_path & Input CSV path.\\
delimiter & Separator (\verb|;| by default, or \verb|,|, tab, or literal text).\\
dependent\_choice & Target variable (\verb|last| by default; accepts column names or positions).\\
non\_variable\_spec & Columns to exclude from the predictor space.\\
exclude\_constant, constant\_threshold & Enable and set the nearly constant predictor filter (percentage).\\
excluded\_observations & Observation IDs or ranges to omit.\\
method & Search method: \texttt{eprs} or \texttt{all\_subsets}.\\
max\_vars & Max predictors per model.\\
n\_seeds, seed\_size & EPR-S seed count and size.\\
iterations\_mode, max\_iterations\_per\_seed & Iterations per seed (\texttt{auto}, \texttt{manual}, or \texttt{converge}); manual requires the max-iterations value.\\
target\_metric & Objective metric (R2, R2 adj, or Q2 (LOO)).\\
cov\_type & Covariance type: Non-robust, HC0, HC1, HC2, or HC3.\\
signif\_lvl & Significance level.\\
corr\_threshold, vif\_threshold & Correlation threshold and VIF threshold for multicollinearity pruning.\\
report\_r2\_threshold & Minimum report R$^2$ (training R$^2$ gate for recording models).\\
clip\_enabled, clip\_low, clip\_high & Toggle and bounds for clipping predictions.\\
split\_mode & \texttt{none} (full dataset), \texttt{random} (test \%), \texttt{manual} (train/test ID lists), or \texttt{external} (validation file).\\
external\_path, external\_delimiter & External validation file path and delimiter (when \texttt{split\_mode = external}).\\
manual\_train\_ids, manual\_test\_ids & ID lists/ranges for manual splits.\\
random\_test\_size\_percent & Test percentage for random splits.\\
export\_limit & Max rows written to the model CSV and linked tables (default 5000).\\
n\_jobs & Parallel jobs (normalized to the allowed maximum).\\
output\_path & Destination CSV for models.\\
\end{longtable}

\paragraph{Minimal example}
\begin{lstlisting}[style=mlrx]
# Dataset
data_path: data.csv
delimiter: ;
dependent_choice: last
non_variable_spec: ID
exclude_constant: true
constant_threshold: 90

# Method
method: eprs
max_vars: 10
n_seeds: 20
seed_size: 4
iterations_mode: auto
signif_lvl: 0.05
corr_threshold: 0.90
vif_threshold: 4.0
report_r2_threshold: 0.80
target_metric: R2
cov_type: HC3
export_limit: 500
n_jobs: 4

# Validation
split_mode: random
random_test_size_percent: 20

# Output
output_path: results/models.csv
clip_enabled: false
\end{lstlisting}

\subsection{Execution and progress}
The CLI loads the configuration, prints the dataset name, available predictors, search method, and (for EPR-S) the seed and iteration settings. It validates ranges (including manual iteration counts and split IDs), normalizes excessive parallel job requests, and streams progress updates for model searching and VIF calculations. Upon completion it reports models explored, models meeting the Minimum report R$^2$ threshold, filtered models kept after export limits, and CPU times.

\subsection{Output}
Results are exported to the specified CSV with semicolon separation, normalized variable names, training/internal/external metrics, and a metadata header so the GUI can reload the run. If no model satisfies the Minimum report R$^2$ filter, no CSV is written. Any parsing or runtime error is printed to stderr and exits with code 1.

% =========================================================
\section{Operational good practices}
\begin{itemize}
  \item Review parameters before each run and save configuration files (\texttt{.conf}) for batch/CLI reproducibility.
  \item Adjust \emph{Parallel jobs} to available hardware to balance time and resources.
  \item Use \emph{Iterations mode = Converge} to let EPR-S control iterations by convergence.
  \item Check collinearity with correlation and VIF thresholds; confirm robustness in \textbf{Validation} and \textbf{Diagnostics} before decisions.
  \item Keep the CSV clean and document exclusions; for external holdout, preserve identical headers and order.
\end{itemize}

% =========================================================
\section{Troubleshooting and FAQ}
\begin{itemize}
  \item \textbf{The GUI does not open}: reinstall the corresponding build and ensure the binary is not blocked by antivirus.
  \item \textbf{CSV does not load}: verify path, permissions, and delimiter; use \emph{Preview data}.
  \item \textbf{No models exported}: review \emph{Minimum report R$^2$} and \emph{VIF threshold}.
  \item \textbf{External validation fails}: the target column must match and the CSV must contain the same predictors.
  \item \textbf{Empty plot or missing labels}: select a valid model/dataset or load a results file.
  \item \textbf{Unexpected CPU usage}: check \emph{Parallel jobs}; the system normalizes the value to at least 1 and reports adjustments.
\end{itemize}

% =========================================================
\section{Credits, license, and citation}
\begin{itemize}
  \item Software distributed under AGPL-3.0. See \texttt{LICENSE.txt} for the full text.
  \item Author and copyright: Jackson J. Alcázar (2025).
  \item Suggested citation: \\ \emph{Alcázar, Jackson J. ``MLR-X 1.0: A Scalable Software for Multiple Linear Regression on Small and Large Datasets''}. \texttt{CITATION\_TEXT} and BibTeX entry \texttt{CITATION\_BIB} are included in the source code.
  \item Contact and improvements: open issues in the official repository; contributions are welcome.
\end{itemize}

% =========================================================
\section{Version history 1.0}
\begin{itemize}
  \item Initial release with full GUI and CLI support.
  \item Inclusion of Linux and Windows builds (GUI and GUI+CLI); macOS build announcement.
  \item Deferred loading of heavy dependencies and automatic parallel execution.
  \item Internal (LOO/k-fold) and external validations; export of summaries and high-resolution plots.
\end{itemize}

\end{document}
